### 2016-06-06 conclusions.tex
Reorganizing the document and adding basic conclusions

\label{sec:conclusions}
Given the need for a \ac{Raspi} testbed easy to configure and maintain
for network coding applications where results could be reproducible
in a simple fashion, we provide a testbed description at Aalborg
University with an in-depth performance evaluation for network coding
schemes to indicate the processing capabilities and energy behaviour
for two \ac{Raspi} models both of using their \ac{SIMD} capabilities
and considering a multicore implementation for the \ac{Raspi} 2. Our
findings report that processing speeds of more than X Mbps and Y Mbps are attainable for \textbf{[goodput scenarios]}, respectively, and that
processing energy per bit for network coding is below 1 nJ in a large regime
and even an order of magnitude lower for some algorithms with the same power consumption for \textbf{[goodput scenarios]}. Future work in this area,
will be to evaluate further elaborated codecs, expand the setup and
automation of tasks within the testbed creation, consider other working
topologies and include a description for network connectivity where the
testbed has more autonomy with \ac{IP} address assignment.