### 2016-06-06 measurements.tex
Average power done. Going for energy

%!TEX root = raspi_journal.tex
\label{sec:measurements}

With the testbed, methodology and setups from the previous sections, we
proceed to obtain the measurements for the \ac{Raspi} 1 and 2 platforms.
We consider the following set of parameters for our study:
For all the codes, we use $g = [16,\ 32,\ 64,\ 128,\ 256,\ 512]$
and $q = [2,\ 2^8]$. For the non-optimized multicore implementations
and cases when the generation size is varied $B = 1600$ bytes.
We consider a setup where we only vary the packet size,
$B = [64,\ 128,\ 256,\ 512,\ 1024,\ 2048,\ 4096,\ 8192,\ 16384,\ 32768
,\ 65536,\ 131072]$ bytes with a generation size fixed on $g = [16,\ 128]$
to see computing load that the \ac{Raspi} present in the low and high
packet size regimes. We considered $B = 1536$ bytes for the optimized
multicore implementation. For \ac{SRLNC}, we considered
$d = [0.02,\ 0.1]$. For \ac{TSNC}, we considered excess packets,
$e = [8,\ 16]$, so that the budget is simply $b = g + e$. In all our
measurement reports, to simplify their review, we first presents the
results for the \ac{Raspi} 1 and later continue with the \ac{Raspi} 2.

\subsection{Goodput}
For the goodput, we separate the results according to their time benchmarks
as we did in Section \ref{sec:metrics} since their consider different
methodologies. We proceed first with the measurements related to the encoder
and later review the case of the decoder.

\subsubsection{Encoding}
Fig. \ref{fig:enc_good_rasp1} shows the results of the encoder goodput
measurements for the \ac{Raspi} 1. Figs. \ref{fig:enc_good_rasp1_gen_gf2}
and \ref{fig:enc_good_rasp1_gen_gf256} present goodput as function of the
generation size for $GF(2)$ and $GF(2^8)$ respectively for a packet size
of $B = 1600$ bytes. Figs. \ref{fig:enc_good_rasp1_packet_gf2} and
\ref{fig:enc_good_rasp1_packet_gf256} present the same metric but now as
function of the packet size for $GF(2)$ and $GF(2^8)$ respectively. In this
case the generation size was set to $g = 16$ packets.

Similarly, Fig. \ref{fig:enc_good_rasp2} shows the same set of
measurements for the \ac{Raspi} 2. Hence, Figs.
\ref{fig:enc_good_rasp2_gen_gf2} and \ref{fig:enc_good_rasp2_gen_gf256}
present goodput as function of the generation size  and
Figs. \ref{fig:enc_good_rasp2_packet_gf2} and
\ref{fig:enc_good_rasp2_packet_gf256} as a function of the packet size
for the same cases mentioned previously. Therefore, we will proceed
to make the results analysis with \ac{Raspi} 1 and indicate which
similarities or differences occur for the \ac{Raspi} 2 and when do
they occur.
%
\begin{figure*}
    \centering
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=1.15\textwidth]{images/06_06_2016/goodput_vs_generation_size_Rasp_encoder_Binary_1600.eps}
        \caption[]%
        {{\small Goodput vs. Generation size for $q = 2$}}
        \label{fig:enc_good_rasp1_gen_gf2}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=1.15\textwidth]{images/06_06_2016/goodput_vs_generation_size_Rasp_encoder_Binary8_1600.eps}
        \caption[]%
        {{\small Goodput vs. Generation size for $q = 2^8$}}
        \label{fig:enc_good_rasp1_gen_gf256}
    \end{subfigure}
    \vskip\baselineskip
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=1.15\textwidth]{images/06_06_2016/goodput_vs_symbol_size_Rasp_encoder_Binary_16.eps}
        \caption[]%
        {{\small Goodput vs. Packet size for $q = 2,\ g=16$}}
        \label{fig:enc_good_rasp1_packet_gf2}
    \end{subfigure}
    \quad
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=1.15\textwidth]{images/06_06_2016/goodput_vs_symbol_size_Rasp_encoder_Binary8_16.eps}
        \caption[]%
        {{\small Goodput vs. Packet size for $q = 2^8,\ g=16$}}
        \label{fig:enc_good_rasp1_packet_gf256}
    \end{subfigure}
    \caption[]
    {\small Encoder goodput measurements for the \ac{Raspi} 1}
    \label{fig:enc_good_rasp1}
\end{figure*}
%
As it can seen in the from: Figs. \ref{fig:enc_good_rasp1_gen_gf2},
\ref{fig:enc_good_rasp1_gen_gf256}, \ref{fig:enc_good_rasp2_gen_gf2} and
\ref{fig:enc_good_rasp2_gen_gf256}; which indicate the goodput dependency
on the generation size, the encoding goodput gets reduced as the generation
size increases, regardless of the \ac{Raspi} model observed. The reason
is that the encoding operation processing is $\mathcal{O}(g)$ because
it is required to create $g$ coding coefficients and do $g$ multiplications
for each packet. This makes the goodput to scale as the inverse of the
generation size.

In the sets of figures shown in Figs.~\ref{fig:enc_good_rasp1} and
\ref{fig:enc_good_rasp2}, it can be observed that the goodput for $GF(2)$
is higher than for $GF(2^8)$ but still around the same order of magnitud.
This difference is explained by noticing that \ac{GF} arithmetics in the
binary field are simply XOR or AND operations. These operations are
implemented efficiently by most of the architectures nowday and are
tailored to be fast in principle. However, the operations in $GF(2^8)$
are more complex given that the finite field arithmetics have to be
performed with lookup tables which at the end reduces the computing
speed giving a lower goodput when compared with the binary field.

In \ref{fig:enc_good_rasp1_gen_gf2}, \ref{fig:enc_good_rasp1_gen_gf256},
\ref{fig:enc_good_rasp2_gen_gf2} and \ref{fig:enc_good_rasp2_gen_gf256};
it can be seen the goodput trends of 5 codes as a function of
the generation size: \ac{RLNC}, \ac{SRLNC} with $d = [0.02,\ 0.1]$
and \ac{TSNC} with an extra parameter that we mention. We define
$C$ as the number of density changes in the \ac{TSNC} transmission
process, if any. The maximum number of possible changes will depend
on the budget. The larger the budget, the more regions are formed and
more density changes are possible. We first evaluate \ac{TSNC} with no
changes, e.g. $C = 0$ and the two amounts for dependent packets mentioned
earlier. This means that we tolerate at most 8 or 16 packets regardless of
the generation. The goal with setting first \ac{TSNC} with these values
is to observe first the sole effect of the budget in the code since
$C = 0$ implies no density changes during the whole process.

For the plots in the mentioned figures from the encoder goodput measurements,
\ac{RLNC} presents the lowest performance in terms of goodput and
\ac{TSNC} with $e = 16$, the highest regardless of the \ac{Raspi} model.
Given that the processing time depends in the amount of coded packets
required to create, \ac{RLNC} is the slowest to process since it must
use all of the $g$ original packets. Later, sparse codes process the data
at a larger rate since less packets are being mixed when creating a
coded packet. The caveat of these schemes is that the sparser the code,
the more probable the ocurrences of \ac{l.d.} packets is. So, basically
the sparser the codes, then more overhead due to transmissions of \ac{l.d.}
packets since it is more frequent to generate similar set of packets.
Excluding the coding coefficients overhead, the overhead due to
transmissions of \ac{l.d.} packets might be high for the sparse schemes.
For example, if we consider \ac{TSNC} with $e = 16$ and $g = 16$
in Fig.~\ref{fig:enc_good_rasp1_gen_gf2}, the budget in this case permits
to send up to 32 packets which is $2\times$ the generation
size for an overhead of $100\%$ excluding the overhead from appending
the coding coefficients. This happens because \ac{TSNC}
has been allowed to add too much redundancy in this case. Given that $C=0$,
this permits it to use a single very low code density $d$. For \ac{RLNC},
this is not the case since the occurrence of \ac{l.d.} coded packets is
small, because all coding coefficients are used. Even for $GF(2)$, the
average amount of redundant packets for \ac{RLNC} has proven to be 1.6
packets after $g$ have been transmitted \textbf{[Include a reference]},
but less than the cases where sparse codes are utilized. Overall, we
observe that there is a trade-off between goodput and \ac{l.d.} coded
packets transmission overhead.

For Figs.~\ref{fig:enc_good_rasp1_packet_gf2},
\ref{fig:enc_good_rasp1_packet_gf256}, \ref{fig:enc_good_rasp2_packet_gf2}
and \ref{fig:enc_good_rasp2_packet_gf256}; we see the packet
size effect in the encoding goodput in both \ac{Raspi} models and for
a fixed generation size. For \ac{TSNC} in this case, we allow for 5 density
changes during the generation transmission and again consider the same
budgets as before. In all the cases, we observe that there is an
optimal packet size to operate with. When reviewing the specifications
of the \ac{Raspi} 1, it uses a L1 cache of 16 KiB. Hence, the trends
in the packet size can be explained in the following way: For low packet
sizes, the data processing does not overload the \ac{CPU} of the
\ac{Raspi}, so the goodput progressively increases given that we simply
process more data as the packet size increases. However, after a certain
packet size, the goodput gets affected since the cache starts to saturate.
The packets towards the cache needs to have more processing. Beyond this
critical packet size, the \ac{CPU} just queues processing which incurs in
larger delay reducing the goodput.

If we consider the previous effect when reviewing the trends in the
mentioned figures, in all the models we observe that the maximum is not
at 16 KiB but at a smaller value depending on the model, field size
and coding scheme considered. The reason is that the \ac{CPU} needs to
allocate computational resources to other processes from the different tasks
running in the \ac{Raspi}. Then, given there are various
tasks from other applications for proper functioning running in the
\ac{Raspi} at the same time as coding, the cache space is filled also
with the data from these tasks, thus diminishing the cache space available
for coding operations.

For the \ac{Raspi} 1 model, we observe in
Fig.~\ref{fig:enc_good_rasp1_packet_gf2} that the critical
packet size for \ac{RLNC} using $GF(2)$ occurs at 1 KiB whereas for the
sparse codes is close to 8 KiB in the most of the cases for $GF(2)$. This
difference takes place since the sparse codes mix less packets their
process than \ac{RLNC}, which turns into less data loading in the
cache for doing computations. For a density of $d = 0.1$, packet size of
$B = 8$ KiB and $g = 16$ packets, we observe that roughly $\ceil{gd} = 2$
packets are inserted in the cache when calculating a coded packet with this
sparse code. Loading this into the cache, stands for 16 KiB which is the
cache size. A similar effect occurs for the other sparse codes. However, for
\ac{RLNC} given that it is a dense code since $d \to 1$, \ac{RLNC}
packets load data from all the $g = 16$ coding coefficients, which accounts
for the 16 KiB of the cache size. In
Fig.~\ref{fig:enc_good_rasp1_packet_gf256}, although the ideal packet
size remains the same for \ac{RLNC} and the sparse codes, the final goodput
is lower due to the field size effect.
%
\begin{figure*}
    \centering
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=1.15\textwidth]{images/06_06_2016/goodput_vs_generation_size_Rasp_v2_encoder_Binary_1600.eps}
        \caption[]%
        {{\small Goodput vs. Generation size for $q = 2$}}
        \label{fig:enc_good_rasp2_gen_gf2}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=1.15\textwidth]{images/06_06_2016/goodput_vs_generation_size_Rasp_v2_encoder_Binary8_1600.eps}
        \caption[]%
        {{\small Goodput vs. Generation size for $q = 2^8$}}
        \label{fig:enc_good_rasp2_gen_gf256}
    \end{subfigure}
    \vskip\baselineskip
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=1.15\textwidth]{images/06_06_2016/goodput_vs_symbol_size_Rasp_v2_encoder_Binary_16.eps}
        \caption[]%
        {{\small Goodput vs. Packet size for $q = 2,\ g=16$}}
        \label{fig:enc_good_rasp2_packet_gf2}
    \end{subfigure}
    \quad
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=1.15\textwidth]{images/06_06_2016/goodput_vs_symbol_size_Rasp_v2_encoder_Binary8_16.eps}
        \caption[]%
        {{\small Goodput vs. Packet size for $q = 2^8,\ g=16$}}
        \label{fig:enc_good_rasp2_packet_gf256}
    \end{subfigure}
    \caption[]
    {\small Encoder goodput measurements for the \ac{Raspi} 2}
    \label{fig:enc_good_rasp2}
\end{figure*}
%
\par
The effects mentioned for the \ac{Raspi} 1 were also observed for
the \ac{Raspi} 2 as mentioned previously. Still, the \ac{Raspi} 2 achieves
roughly $5\times$ to $7\times$ gains in terms of encoding speed when comparing
the goodputs in Figs.~\ref{fig:enc_good_rasp1} and \ref{fig:enc_good_rasp2},
given that it has an \ac{ARM} Cortex A7 (v7) \ac{CPU} and twice the \ac{RAM}
size than the \ac{ARM}1176JZF-S (v6) core of the \ac{Raspi} 1 model.

In Figs.\ref{fig:enc_good_rasp2_packet_gf2} and
\ref{fig:enc_good_rasp2_packet_gf256}, we observe that the packet size for
the maximum \ac{RLNC} goodput has shifted towards 8 KiB, indicating that
the \ac{Raspi} 2 is able to handle $16 \times 8$ KiB $= 128$ KiB. This is
possible because the \ac{Raspi} 2 has a shared L2 cache of 256 KiB allowing
to still allocate some space for the data to be processed while achieving
a maximum goodputs of 105 MB/s.

\subsubsection{Decoding}
Similar as the encoding goodput, in this Section, we review decoder
goodput in terms of performance and configurations.
Fig.~\ref{fig:dec_good_rasp1} shows the results of the decoder goodput
measurements for the \ac{Raspi} 1 and Fig. \ref{fig:dec_good_rasp2} does
the proper for \ac{Raspi} 2.

\begin{figure*}
    \centering
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=1.15\textwidth]{images/06_06_2016/goodput_vs_generation_size_Rasp_decoder_Binary_1600.eps}
        \caption[]%
        {{\small Goodput vs. Generation size for $q = 2$}}
        \label{fig:dec_good_rasp1_gen_gf2}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=1.15\textwidth]{images/06_06_2016/goodput_vs_generation_size_Rasp_decoder_Binary8_1600.eps}
        \caption[]%
        {{\small Goodput vs. Generation size for $q = 2^8$}}
        \label{fig:dec_good_rasp1_gen_gf256}
    \end{subfigure}
    \vskip\baselineskip
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=1.15\textwidth]{images/06_06_2016/goodput_vs_symbol_size_Rasp_decoder_Binary_16.eps}
        \caption[]%
        {{\small Goodput vs. Packet size for $q = 2,\ g=16$}}
        \label{fig:dec_good_rasp1_packet_gf2}
    \end{subfigure}
    \quad
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=1.15\textwidth]{images/06_06_2016/goodput_vs_symbol_size_Rasp_decoder_Binary8_16.eps}
        \caption[]%
        {{\small Goodput vs. Packet size for $q = 2^8,\ g=16$}}
        \label{fig:dec_good_rasp1_packet_gf256}
    \end{subfigure}
    \caption[]
    {\small Decoder goodput measurements for the \ac{Raspi} 1}
    \label{fig:dec_good_rasp1}
\end{figure*}

In Figs.~\ref{fig:dec_good_rasp1} and \ref{fig:dec_good_rasp2}, we
observe the same generation and packet size effects reported in the
encoding case. However, we do observe in
Figs.~\ref{fig:dec_good_rasp1_gen_gf2}, \ref{fig:dec_good_rasp1_gen_gf256},
\ref{fig:dec_good_rasp2_gen_gf2} and \ref{fig:dec_good_rasp2_gen_gf256};
that duplicating the generation size does not reduce by a factor of 4 the
goodput. In principle, given that Gaussian elimination scales as
$\mathcal{O}(g^3)$ (and thus the processing time), we would expect that
the goodput scales as $\mathcal{O}(R_{proc,dec}) = \mathcal{O}(g/g^{3}) =
\mathcal{O}(g^{-2})$. This would imply that duplicating the generation
size, should reduce the goodput by a factor of 4, which does not. Instead,
the goodput it is only reduced by a factor of 2. This is only possible if
the Gaussian elimination is $\mathcal{O}(g^2)$. A study in
\textbf{[refence about Achutan's measurements on the Raspi]} indicated
that this is effectively the case. The reason is that the $g^2$ factor in
the scaling law for our Gaussian elimination is much higher than the $g^3$
for $g \geq 512$. \textbf{[Why was the $g^2$ factor greather than the
$g^3$ factor? Wasn't it due to the amount of of bytes per packet?]}

Another difference with the encoding goodput results in the same figures,
is that even though the \ac{TSNC} with $e = 16$ packets provides the fastest
encoding, it does not happen to be the same for the decoding. For this
very sparse scheme, the decoding is affected by the amount of \ac{l.d.}
packets generated by, which leads to a higher delay in some cases
(particularly in $g = [64,\ 128]$). For other generation sizes, the
performance of sparse codes is similar.
%
\begin{figure*}
    \centering
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=1.15\textwidth]{images/06_06_2016/goodput_vs_generation_size_Rasp_v2_decoder_Binary_1600.eps}
        \caption[]%
        {{\small Goodput vs. Generation size for $q = 2$}}
        \label{fig:dec_good_rasp2_gen_gf2}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=1.15\textwidth]{images/06_06_2016/goodput_vs_generation_size_Rasp_v2_decoder_Binary8_1600.eps}
        \caption[]%
        {{\small Goodput vs. Generation size for $q = 2^8$}}
        \label{fig:dec_good_rasp2_gen_gf256}
    \end{subfigure}
    \vskip\baselineskip
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=1.15\textwidth]{images/06_06_2016/goodput_vs_symbol_size_Rasp_v2_decoder_Binary_16.eps}
        \caption[]%
        {{\small Goodput vs. Packet size for $q = 2,\ g=16$}}
        \label{fig:dec_good_rasp2_packet_gf2}
    \end{subfigure}
    \quad
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=1.15\textwidth]{images/06_06_2016/goodput_vs_symbol_size_Rasp_v2_decoder_Binary8_16.eps}
        \caption[]%
        {{\small Goodput vs. Packet size for $q = 2^8,\ g=16$}}
        \label{fig:dec_good_rasp2_packet_gf256}
    \end{subfigure}
    \caption[]
    {\small Decoder goodput measurements for the \ac{Raspi} 2}
    \label{fig:dec_good_rasp2}
\end{figure*}
%
\subsection{Average Power}
With the energy measurement setup described in Section
\ref{sec:energy_metrics_methods}, we compute the average power of each
device from their current consumption across time in
Figs.~\ref{fig:current_rasp1} and \ref{fig:current_rasp2}. In each
figure, the eletrical current values are classified into three possible
currents: Idle, transition and processing. These values are identified by
our post-processing. Later, the values in red are idle currents, green for
transitions and blue for processing.

For a given configuration of a coding scheme, device type and its parameters,
only electrical current processing values are present. In general, we
associate the current samples from a given configuration with its goodput
through a timestamp. In this way, we relate goodput and power / energy
measurements. Once identified the current values, we compute the averages
for each configuration as described in Section
\ref{sec:energy_metrics_methods}. In the presented current samples, we observe
a presence of bursty noise. Nevertheless, by taking the average as described
in \eqref{eq:avg_current}, we remove this contribution in the average
processing current.
%
\begin{figure*}
    \centering
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=1.1\textwidth]{images/06_06_2016/current_vs_sec_raspberry_change_gen_from_-10sec_to_500sec.eps}
        \caption[]%
        {{\small Current for the \ac{Raspi} 1 model}}
        \label{fig:current_rasp1}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=1.1\textwidth]{images/06_06_2016/current_vs_sec_raspberry2_change_gen_from_-10sec_to_500sec.eps}
        \caption[]%
        {{\small Current for the \ac{Raspi} 2 model}}
        \label{fig:current_rasp2}
    \end{subfigure}

    \caption[]
    {\small Current consumption for each \ac{Raspi} model}
    \label{fig:current_rasp}
\end{figure*}
%
By reviewing the processing current samples in Figs.~\ref{fig:current_rasp1}
and \ref{fig:current_rasp2}, we observe that the average processing current
does not change significantly for each \ac{Raspi} model in all the shown
configurations. Therefore, we approximate the current consumption
for the devices and compute the average power as indicated in
\eqref{eq:avg_power}. The results are shown in
Table~\ref{tab:average_power_table}.

\begin{table}[H]
\center
\caption{Average power for the \ac{Raspi} models}
\begin{tabular}{|r|r|r|r|}

\hline
\ac{Raspi} Model & $I_{idle,avg}$ (A)& $I_{proc,avg}$ (A) & $P_{avg}$ (W) \\
\hline
\hline
    \ac{Raspi} 1 &      0.320        &            0.360    & 0.200\\
\hline
    \ac{Raspi} 2 &      0.216        &            0.285    & 0.345\\
\hline
\end{tabular}
\vspace{0.2cm}
\label{tab:average_power_table}
\end{table}
%
From Table~\ref{tab:average_power_table}, we notice that the power expenditure
for both models is almost the same. Thus, the energy behaviour is mostly
dependent on the goodput trends since the power is just an scaling constant.
%
\subsection{Energy per bit}
With power and goodput measurements, we compute the energy per bit of
the previously mentioned cases as described in \eqref{eq:energy_per_bit}.
The trends of the energy per bit are simply the inverse of the goodput since
both metrics are related by an inverse law.

\subsubsection{Encoding}

\begin{figure*}
    \centering
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=1.1\textwidth]{images/06_06_2016/energy_per_bit_vs_generation_size_Rasp_Binary_encoder_1600.eps}
        \caption[]%
        {{\small Energy per bit vs. Generation size for $q = 2$}}
        \label{fig:enc_ene_rasp1_gen_gf2}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=1.1\textwidth]{images/06_06_2016/energy_per_bit_vs_generation_size_Rasp_Binary8_encoder_1600.eps}
        \caption[]%
        {{\small Energy per bit vs. Generation size for $q = 2^8$}}
        \label{fig:enc_ene_rasp1_gen_gf256}
    \end{subfigure}
    \vskip\baselineskip
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=1.1\textwidth]{images/06_06_2016/energy_per_bit_vs_symbol_size_Rasp_Binary_encoder_128.eps}
        \caption[]%
        {{\small Energy per bit vs. Packet size for $q=2,\ g=128$}}
        \label{fig:enc_ene_rasp1_packet_gf2}
    \end{subfigure}
    \quad
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=1.1\textwidth]{images/06_06_2016/energy_per_bit_vs_symbol_size_Rasp_Binary8_encoder_128.eps}
        \caption[]%
        {{\small Energy per bit vs. Packet size for $q=2^8,\ g=128$}}
        \label{fig:enc_ene_rasp1_packet_gf256}
    \end{subfigure}
    \caption[]
    {\small Encoder energy measurements for the \ac{Raspi} 1}
    \label{fig:enc_ene_rasp1}
\end{figure*}

Raspi 2 encoding

\begin{figure*}
    \centering
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=1.1\textwidth]{images/06_06_2016/energy_per_bit_vs_generation_size_Rasp_v2_Binary_encoder_1600.eps}
        \caption[]%
        {{\small Energy per bit vs. Generation size for $q = 2$}}
        \label{fig:enc_ene_rasp2_gen_gf2}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=1.1\textwidth]{images/06_06_2016/energy_per_bit_vs_generation_size_Rasp_v2_Binary8_encoder_1600.eps}
        \caption[]%
        {{\small Energy per bit vs. Generation size for $q = 2^8$}}
        \label{fig:enc_ene_rasp2_gen_gf256}
    \end{subfigure}
    \vskip\baselineskip
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=1.1\textwidth]{images/06_06_2016/energy_per_bit_vs_symbol_size_Rasp_v2_Binary_encoder_128.eps}
        \caption[]%
        {{\small Energy per bit vs. Packet size for $q=2,\ g=128$}}
        \label{fig:enc_ene_rasp2_packet_gf2}
    \end{subfigure}
    \quad
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=1.1\textwidth]{images/06_06_2016/energy_per_bit_vs_symbol_size_Rasp_v2_Binary8_encoder_128.eps}
        \caption[]%
        {{\small Energy per bit vs. Packet size for $q=2^8,\ g=128$}}
        \label{fig:enc_ene_rasp2_packet_gf256}
    \end{subfigure}
    \caption[]
    {\small Encoder energy measurements for the \ac{Raspi} 2}
    \label{fig:enc_ene_rasp2}
\end{figure*}

\subsubsection{Decoding}

\begin{figure*}
    \centering
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=1.1\textwidth]{images/06_06_2016/energy_per_bit_vs_generation_size_Rasp_Binary_decoder_1600.eps}
        \caption[]%
        {{\small Energy per bit vs. Generation size for $q = 2$}}
        \label{fig:dec_ene_rasp1_gen_gf2}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=1.1\textwidth]{images/06_06_2016/energy_per_bit_vs_generation_size_Rasp_Binary8_decoder_1600.eps}
        \caption[]%
        {{\small Energy per bit vs. Generation size for $q = 2^8$}}
        \label{fig:dec_ene_rasp1_gen_gf256}
    \end{subfigure}
    \vskip\baselineskip
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=1.1\textwidth]{images/06_06_2016/energy_per_bit_vs_symbol_size_Rasp_Binary_decoder_128.eps}
        \caption[]%
        {{\small Energy per bit vs. Packet size for $q=2,\ g=128$}}
        \label{fig:dec_ene_rasp1_packet_gf2}
    \end{subfigure}
    \quad
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=1.1\textwidth]{images/06_06_2016/energy_per_bit_vs_symbol_size_Rasp_Binary8_decoder_128.eps}
        \caption[]%
        {{\small Energy per bit vs. Packet size for $q=2^8,\ g=128$}}
        \label{fig:dec_ene_rasp1_packet_gf256}
    \end{subfigure}
    \caption[]
    {\small Decoder energy measurements for the \ac{Raspi} 1}
    \label{fig:dec_ene_rasp1}
\end{figure*}

Raspi 2 decoding

\begin{figure*}
    \centering
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=1.1\textwidth]{images/06_06_2016/energy_per_bit_vs_generation_size_Rasp_v2_Binary_decoder_1600.eps}
        \caption[]%
        {{\small Energy per bit vs. Generation size for $q = 2$}}
        \label{fig:dec_ene_rasp2_gen_gf2}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=1.1\textwidth]{images/06_06_2016/energy_per_bit_vs_generation_size_Rasp_v2_Binary8_decoder_1600.eps}
        \caption[]%
        {{\small Energy per bit vs. Generation size for $q = 2^8$}}
        \label{fig:dec_ene_rasp2_gen_gf256}
    \end{subfigure}
    \vskip\baselineskip
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=1.1\textwidth]{images/06_06_2016/energy_per_bit_vs_symbol_size_Rasp_v2_Binary_decoder_128.eps}
        \caption[]%
        {{\small Energy per bit vs. Packet size for $q=2,\ g=128$}}
        \label{fig:dec_ene_rasp2_packet_gf2}
    \end{subfigure}
    \quad
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=1.1\textwidth]{images/06_06_2016/energy_per_bit_vs_symbol_size_Rasp_v2_Binary8_decoder_128.eps}
        \caption[]%
        {{\small Energy per bit vs. Packet size for $q=2^8,\ g=128$}}
        \label{fig:dec_ene_rasp2_packet_gf256}
    \end{subfigure}
    \caption[]
    {\small Decoder energy measurements for the \ac{Raspi} 2}
    \label{fig:dec_ene_rasp2}
\end{figure*}

\subsection{Multicore Network Coding}
\label{subs:multicore-network-coding}

The authors implemented the algorithm described in
Section~\ref{sub:implementation-multicore} on the Raspberry Pi 2 Model B,
which features four ARM Cortex-A7 cores in a Broadcom BCM2836 \ac{SOC}
with a 900 MHz clock. Each core has a 32 KiB L1 data cache and 32 KiB
L1 instruction cache. The cores share a 512 KiB L2 cache. All the
measured results, including the baseline results, were obtained with
NEON-enabled code adopted from the Fifi library [\textbf{Fifi reference}].
The NEON extension provides an 128-bit \ac{SIMD} instruction set to the
\ac{Raspi} 2. Figures~\ref{enc_dec1024}, \ref{enc_dec128} and
\ref{enc_dec16} show the encoding and decoding throughput in MiB per
second for different generation sizes ($g$ = 1024, 128, and 16
respectively). The throughput is the rate of generating $g$ coded packets,
divided by the time the encoder took to perform the task. The size of
each coded packet was fixed to $1536$~bytes since that is the typical
size of an Ethernet frame. The blocked operations were performed dividing
the matrices in squared sub-blocks of $16,\ 32\ ,\ 64,\ldots,\ 1024$ operands
(words in the finite field) in height and width. The figures show only
block sizes of $16 \times 16$, $32 \times 32$, $64 \times 64$ operands
since with bigger block sizes, the operands do not fit in L1 cache.

The following test cases were considered:

\subsubsection{Baseline encoding}
The baseline results involve no recording of the
\ac{DAG} and are performed in a \emph{by-the-book} fashion. The encoder uses
only one thread. And the difference between the non-blocked and blocked
encoding schemes is that in the blocked scheme, the matrix multiplications are
performed dividing the matrices in sub-blocks in order to make the algorithm
cache efficient as described in Section~\ref{sub:implementation-multicore}.

\subsubsection{Encoding blocked}
The encoding results were performed using the
method described in Section~\ref{sub:implementation-multicore}. The time
recorded includes the dependencies resolving, creation of the \ac{DAG},
and the task scheduling. In practice, it would suffice to calculate and
store this information only once per generation size.

\subsubsection{Decoding blocked}
The differences between encoding and decoding, is that the decoding task
also involves the matrix inversion. Similarly as with the encoding results,
the time recorded includes the dependencies resolving, the creation of the
\ac{DAG} and the task scheduling. However, to decode, these calculations are also made for inverting the matrix of coding coefficients.

\begin{figure}[ht!]
\centering
\includegraphics[width=0.7\textwidth]{images/2015-04-18_encoding_decoding_1024.pdf}
\caption{Encoding and Decoding performance for $g = 1024$. Block size: $32 \times 32$. \cite{wunderlich2015network}}
\label{enc_dec1024}
\end{figure}

For $g= 1024$, the blocked baseline measurements outperforms the non blocked
variant. This means that making the matrix multiplication algorithm cache
efficient brings an increase in throughput by a factor of 2.83. When using the
algorithm described in Section~\ref{sub:implementation-multicore},
encoding with four cores is on average 3.7 times faster than with one core.
Similarly, decoding with four codes is 3.05 times faster, on average, than
decoding with a single core. Figure~\ref{enc_dec1024} shows that the
implemented algorithm, by exploiting cache efficiency and only three extra
cores provides a gain of 10 folds compared with traditional non-blocked
algorithms. With $g = 1024$, the matrix inversion becomes more expensive
than at smaller generations sizes. Therefore, the decoding throughput
is 46\% of the encoding throughput.

\begin{figure}[ht!]
\centering
\includegraphics[width=0.7\textwidth]{images/2015-04-18_encoding_decoding_128.pdf}
\caption{Encoding and Decoding performance for $g = 128$. Block size: $64 \times 64$ \cite{wunderlich2015network}}
\label{enc_dec128}
\end{figure}

For $g$ = 128, the differences between the baselines operations show that a
blocked algorithm is 15\% faster than the non-blocked variant. Encoding with
four cores is 2.82 times faster than with a single core. Due to the smaller
matrix sizes, the gain when using blocked operations in the baselines is not
that significant when compared with $g = 1024$. For the same reason, the matrix
inversion is less expensive. As a consequence, the decoding throughput is 71\%
of the encoding throughput.

\begin{figure}[ht!]
\centering
\includegraphics[width=0.7\textwidth]{images/2015-04-18_encoding_decoding_16.pdf}
\caption{Encoding and Decoding performance for $g = 16$. Block size: $16 \times 16$ \cite{wunderlich2015network}}
\label{enc_dec16}
\end{figure}

When $g = 16$, the gains of blocked operations are negligible compared with
the non-blocked ones. The reason behind this behavior is that all the data
fits in the L1 cache. For the scheduled version, since the problem to solve
is so small, the gain when using four cores is a factor of 1.68 compared
with a single core, and 1.01 compared with two cores. Therefore, there
are no practical benefits in using four cores instead of two.

The differences in throughput, for all generation sizes, between the
blocked baseline and the single threaded scheduled measurements are due the
time spent resolving the dependencies and the scheduling overhead. These
effects are negligible for big generation sizes, while considerable for
small matrices.

\subsubsection{Comparison of the load of matrix multiplications and inversions}

To compare how much slower is the matrix multiplication with respect to the
matrix inversion for different generation sizes, we ran a set of tests. We used
a single core to perform the operations. We changed the generation sizes,
performed matrices multiplications and matrix inversions, and measured the time
spent doing so which we name $T_{mult}$ and $T_{inv}$. We calculate the ratio
between these two measured times defined as $r = \frac{T_{mult}}{T_{inv}}$.
Table~\ref{runtimes} summarizes the results. The bigger the matrix size, the
smaller is the calculated ratio. This means that when the problems are bigger,
the decoding throughput decreases compared with the encoding throughput.

\begin{table}[H]
\center
\caption{Multiplication and inversion run-times for different generation sizes with 1 thread}
\begin{tabular}{|r|r|r|r|}

\hline
$g$ & $T_{mult}$ (ms) & $T_{inv}$ (ms) &$r$ \\
\hline
\hline

	16   & 1.703     & 0.345    & 4.9 \\
\hline
	32   & 6.573     & 0.914    & 7.2 \\
\hline
	64   & 21.341    & 3.479    & 6.1 \\
\hline
	128  & 82.326    & 17.411   & 4.7 \\
\hline
	256  & 336.398   & 106.861  & 3.1 \\
\hline
	512  & 1548.750  & 659.469  & 2.3 \\
\hline
	1024 & 6730.380  & 5166.920 & 1.3 \\
\hline
\end{tabular}
\vspace{0.2cm}
\label{runtimes}
\end{table}