### 2016-08-16 reviewer_reply.txt
The remaining work

- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - 
Reviewer 1:
- - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

Dear reviewer,
Thank you very much for your valuable feedback. We have corrected all the indicated and other
found typos by proof-reading the manuscript. We have also replaced the abbreviations "l.d." and
"l.i." to "linearly dependent" and "linearly independent" as suggested to enhance clarity in the 
notation.

FIXME: Nestor, what do you think about the way I defined "g" on page 3? I believe "g" is first used on page 3 now. Not page 2 anymore as the reviewer stated?

Nestor: Where is this exactly right now? I cannot see it :(

FIXME: the notation [1,g] (and many other notations) should be changed to
       maybe [1,2,...g]
Nestor: Can you make this modification while I advance in the intro?

FIXME: change notation such as $\mathbb{N}$ as this is typically used for positive numbers
Nestor: Can you make this modification while I advance in the intro?

- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - 
Reviewer 2:
- - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

Dear reviewer,
Thank you very much for the valuable feedback. We have listed your comments below and described what we have changed to hopefully improve the indicated issues.

 - The proposed system is not well described

We have rewritten the parts that indicate a proposed system. All simulations are performed within a single Raspberry Pi. We only replace the traditional power supply with an Agilent to monitor and log the power consumption.

- Finally and in a word, the motivation of this paper is unclear. The main question is “why Raspberry pi in network coding?”. What is the advantage of using Raspberry pi?

We consider the Raspberry Pi 1 and 2 to be good representatives in the field of low budget and internet of things devices in terms of performance and capabilities. The two devices have one and four cores respectively and they support hardware acceleration that is mandatory to reach the full potential of network coding. We have updated the introduction to hopefully elaborate a bit more on why network coding in Raspberry Pi is interesting. Another reason for using the Raspberry Pi is its large user community in contrast to many other low cost devices.

- They only consider q=2 and 28 , however, as it is mentioned in “Chou, Philip A., Yunnan Wu, and Kamal Jain. "Practical network coding." (2003).”, based on security aspect, in practice, wey should consider q=28 and 216. By this assumption, they should show and compare the results with q=216 too.

We have made a reference to two papers by Achuthan et al. ("Lean and Mean" and "Sharing the Pi" that include measurements for both q=2^8 and q=2^16 that indicates a performance degradation of approximately 10x). We also ran some simulations ourself that indicated similar performance gab between the two field types.
The reason that we do not consider q=2^16 is that nobody uses it today. q=2^8 is the standard today and q=2 is beneficial in terms of performance in low performance devices. For security purposes, there exists better ways than using q=2^16.

FIXME: Did we include achuthans papers?
FIXME: Should we argue somewhere the performance of 2^16 and argue why we use the two fields q=2 and q=2^16?

- The list of references neglects some of the most relevant papers. Moreover, the authors have cited to their previous papers (i.e., [6,7,8,10,11,12,13,17,21,27,30])!!! It seems they want to increase citation of their papers!

We link to our prior work as this paper is mostly based on those papers. We have removed reference 10 and 30 and added a few more citations to related work done by others.

FIXME: Did we remove reference 10 and 30?

Nestor: Which were these papers? These numbers have changed already

- The presentation of the paper is not very good. The authors did not care very much about the structure, the language (i.e., page 2 line 42 “Furthermore, in [7], the Kodo library was introduce.” Should be introduced….)

We have corrected all the mentioned and other found typos by proof-reading the manuscript.

- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - 
Reviewer 3:
- - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

Dear reviewer,
Thank you very much for your valuable feedback. We have corrected all the mentioned and other found
typos as suggested by proof-reading the manuscript and changed the figures in section 4.4 to show the goodput instead of the throughput by properly labelling the axis and verifying that the shown
data is correct.