### 2016-08-16 reviewer_reply.txt
New introduction

- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - 
Reviewer 1:
- - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

Dear reviewer,
Thank you very much for your valuable feedback. We have corrected all the indicated and other
found typos by proof-reading the manuscript. We have also replaced the abbreviations "l.d." and
"l.i." to "linearly dependent" and "linearly independent" as suggested to enhance clarity in the 
notation. Regarding the definition of "g", We have defined the parameter "g" as the total number of original packets to be encoded.

FIXME: the notation [1,g] (and many other notations) should be changed to
       maybe [1,2,...g]
Nestor: Can you make this modification while I advance in the intro?

FIXME: change notation such as $\mathbb{N}$ as this is typically used for positive numbers
Nestor: Can you make this modification while I advance in the intro?

- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - 
Reviewer 2:
- - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

Dear reviewer,
Thank you very much for your valuable feedback. We have listed your comments below and described what we have changed to hopefully improve the indicated issues.

 - The proposed system is not well described

We have rewritten the parts that indicate a proposed system. All our measurements are performed within a single Raspberry Pi. We only replace the traditional power supply with an Agilent Power
Supply device to monitor and log the power consumption.

- Finally and in a word, the motivation of this paper is unclear. The main question is “why Raspberry pi in network coding?”. What is the advantage of using Raspberry pi?

We consider that the Raspberry Pi models 1 and 2 are good representatives of a low-cost computer
considered as IoT devices in terms of performance and capabilities. We now elaborate on this idea in the introduction by providing various use cases of IoT + Raspi. At the end of the description of
the use cases of IoT + Raspi, we specifically mention that despite this advances, data is still
being conveyed in a conventional way that will hit a performance roof with new services demands. This is why we motivate about network coding since it is proven its utility in the area of meshed
networks where IoT deployments occur. Another reason for using the Raspberry Pi is its large user community in contrast to many other low cost devices. This allows a system developer to use standard tools and / or based on either Windows or Linux, reducing the development time. Finally, the Raspberry Pi models 1 and 2, have one and four cores respectively and they support hardware acceleration through the NEON instruction set in ARM that is mandatory to reach the full potential of network coding where we have included the proper details. We have updated the introduction to hopefully elaborate on all these subjects to show why using network coding in Raspberry Pi is interesting.

- They only consider q=2 and 28 , however, as it is mentioned in “Chou, Philip A., Yunnan Wu, and Kamal Jain. "Practical network coding." (2003).”, based on security aspect, in practice, wey should consider q=28 and 216. By this assumption, they should show and compare the results with q=216 too.

We have made a reference to two papers by Achuthan et al. ("Lean and Mean" and "Sharing the Pi").
These two references include measurements for both q=2^8 and q=2^16 and indicate a performance degradation of approximately 10x. The underlying reason is that using Galois fields higher than 2^8 provide no perceivable benefit in terms of the decoding probability while increasing in complexity. At the end, this turns disastrous for the code. We also ran a few measurements ourself that
indicated similar performance gab between the two field types. Another reason to not consider q=2^16 is that it is not a standard field size. q=2^8 is the standard today and q=2 is beneficial in terms of performance in low-computing power devices. For security purposes, there exists even better ways than using q=2^16.

- The list of references neglects some of the most relevant papers. Moreover, the authors have cited to their previous papers (i.e., [6,7,8,10,11,12,13,17,21,27,30])!!! It seems they want to increase citation of their papers!

We link to our prior work as this paper is mostly based on those papers. We have removed reference 10 and 30 and added a few more citations to related work done by others.

FIXME: Did we remove reference 10 and 30?

Nestor: Which were these papers? These numbers have changed already

- The presentation of the paper is not very good. The authors did not care very much about the structure, the language (i.e., page 2 line 42 “Furthermore, in [7], the Kodo library was introduce.” Should be introduced….)

We have corrected all the mentioned and other found typos by proof-reading the manuscript.

- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - 
Reviewer 3:
- - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

Dear reviewer,
Thank you very much for your valuable feedback. We have corrected all the mentioned and other found
typos as suggested by proof-reading the manuscript and changed the figures in section 4.4 to show the goodput instead of the throughput by properly labelling the axis and verifying that the shown
data is correct.