### 2016-08-16 reviewer_reply.txt
Corrected version. Update review reply. Update readme.rst

- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - 
Reviewer 1:
- - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

Dear reviewer,
Thank you very much for your valuable feedback. We have listed your comments below and described
our corrections to hopefully solve the indicated issues.

- We have corrected all the indicated and other found typos by proof-reading the manuscript. We
have also replaced the abbreviations "l.d." and "l.i." to "linearly dependent" and "linearly
independent" as suggested to enhance clarity in the notation.

- Regarding the definition of "g", We have defined the parameter "g" as the total number of
original packets to be encoded once at beginning of Section 2.

- For the notation [1,g] to denote the set of integer number between 1 and "g", we have modified
all its appearances to [1,2,...,g] to indicate the integer nature of this set in a more standard
fashion. Also, in our
definitions the notation

- For the notation $\mathbb{N}$ for the random number of packets to be combined in SRLNC, we have
modified all its appearances to $\mathbb{M}$ to avoid any confusion with the notation for the set of
positive numbers.

- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - 
Reviewer 2:
- - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

Dear reviewer,
Thank you very much for your valuable feedback. We have listed your comments below and described
our corrections or observations to hopefully solve the indicated issues.

- The proposed system is not well described

We have rewritten the parts that indicate a proposed system. All our measurements are performed
within a single Raspberry Pi. We only replace the traditional power supply with an Agilent Power
Supply device to monitor and log the power consumption.

- Finally and in a word, the motivation of this paper is unclear. The main question is "why
Raspberry pi in network coding?". What is the advantage of using Raspberry pi?

We consider that the Raspberry Pi models 1 and 2 are good representatives of a low-cost computer
considered as IoT devices in terms of performance and capabilities. We now elaborate on this idea
in the introduction by providing various use cases of IoT + Raspi. At the end of the description of
the use cases of IoT + Raspi, we specifically mention that despite this advances, data is still
being conveyed in a conventional way that will hit a performance roof with new services demands.
This is why we motivate about network coding, e.g. since the usefulness of network coding has been
verified in the area of meshed networks in prior works. Meshed networks will be the scenario where
upcoming IoT deployments will mostly occur.

Another two important reasons for using the Raspberry Pi are: First, its large user community in
contrast to many other low-cost devices. This allows a system developer to use standard tools
and / or based on either Windows or Linux, reducing the development time. Second, the Raspberry
Pi models 1 and 2, have one and four cores respectively and they support hardware acceleration
through the NEON instruction set in ARM that is mandatory to reach the full potential of network
coding where we have included the proper details. We have updated the introduction to elaborate
on all these subjects to show why using network coding in Raspberry Pi is interesting.

- They only consider q=2 and 28 , however, as it is mentioned in "Chou, Philip A., Yunnan Wu, and
Kamal Jain. "Practical network coding." (2003).", based on security aspect, in practice, wey
should consider q=28 and 216. By this assumption, they should show and compare the results with
q=216 too.

We have made a reference to two papers by Achuthan et al. (i.e. "Lean and Mean" [23] and "Sharing
the Pi" [25]). These two references include measurements for both q=2^8 and q=2^16 and indicate
a performance degradation of approximately 10X for going from q=2^8 to q=2^16. The underlying
reason is that using Galois fields higher than 2^8 provide no perceivable benefit in terms of the
decoding probability while increasing in complexity and processing overhead due to the coding
coefficients. At the end, this significantly reduces the performance of the code. To verify this,
we also ran a few measurements ourself that indicated a similar performance gap between the two
field types. Another reason to not consider q=2^16 is that it is not a standard field size. q=2^8
is the standard today (comparable with Reed-Solomon codes for example) and q=2 is beneficial in
terms of performance in low-computing power devices. Finally, for security purposes, there exists
much secure ways than only using q=2^16.

- The list of references neglects some of the most relevant papers. Moreover, the authors have
cited to their previous papers (i.e., [6,7,8,10,11,12,13,17,21,27,30])!!! It seems they want to
increase citation of their papers!

We have updated the list of references to provide a richer balance in the State-of-the-Art,
relevant baseline references and our prior work.

We now have included several references to previous State-of-the-Art and references related
to use cases for IoT employing the Raspberry Pi [1-9]. Also we have included missing
references and grouped them with the baseline of network coding [10-16,26].

We link to our prior work as this paper evaluation is based on research lines started
in those papers. These prior works report our findings in the area of network coding for
wireless mesh networks and network coding for the Raspberry Pi. To be more specific: Theoretical
gains and practical implementations of network coding for wireless
meshed networks are reported in [17,19,21,22]. The Kodo library, our basic building block for the coding operations is [18]. Basic measurements + observed effects of Random
Linear Network Coding for commercial devices are found in [23]. A initial study of RLNC for
the Raspberry Pi is available in [25]. Finally, our work with Sparse and Tunable Sparse Network
Coding is described in [28-31].

At the end, our references consist of less than 27% of the references in the study. We hope
that this solves the raised issue.

- The presentation of the paper is not very good. The authors did not care very much about the
structure, the language (i.e., page 2 line 42 "Furthermore, in [7], the Kodo library was
introduce." Should be introduced....)

We have corrected all the mentioned and other found typos by proof-reading the manuscript.

- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - 
Reviewer 3:
- - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

Dear reviewer,
Thank you very much for your valuable feedback. We have corrected all the mentioned and other found
typos as suggested by proof-reading the manuscript and changed the figures in section 4.4 to show
the goodput instead of the throughput by properly labelling the axis and verifying that the shown
data is correct.